---
title: "Reproducible research - course project 1"
author: "daniel"
date: "19 March 2018"
output: 
  html_document: 
    keep_md: yes
---

#loading and preprocessing data 

##initialise the environment 
```{r setup, include=TRUE}
library(ggplot2)
setwd("C:/datasciencecoursera/ReproducibleResearche/CourseProject1")
```


##load and proprocessing data

```{r load data, include=TRUE}
dActivity <- read.csv("activity.csv")
#set date format 
dActivity$date <- as.Date(dActivity$date)
dActivity$interval2 <- paste(as.integer(dActivity$interval / 100), formatC(dActivity$interval %% 100, width = 2, flag = 0), sep =":")
```


#Q1 What is mean total number of steps taken per day?

##Q1.1 Calculate the total number of steps taken per day
```{r}
dStepsPerDay <- aggregate(steps~date, data = dActivity, FUN = sum, na.rm = TRUE)
dStepsPerDay
```

##Q1.2 Make a histogram of the total number of steps taken each day
```{r}
barplot(dStepsPerDay$steps, 
        names.arg = dStepsPerDay$date, 
        main = "total step per day", 
        ylab = "# steps", 
        xlab = "date")
```

##Q1.3 Calculate and report the mean and median of the total number of steps taken per day
```{r}
dMeanStepsPerDay   <- aggregate(steps~date, data = dActivity, FUN = mean)
dMedianStepsPerDay <- aggregate(steps~date, data = dActivity, FUN = median)
dMeanMedStepsPerDay <- merge(dMeanStepsPerDay, dMedianStepsPerDay, by = "date")
colnames(dMeanMedStepsPerDay) <- c("date", "mean", "median")
dMeanMedStepsPerDay
```




# hat is the average daily activity pattern?
## Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
dStepsPerInterval <- aggregate(steps~interval, data = dActivity, FUN = mean, na.rm = TRUE)
dStepsPerInterval$interval2 <- paste(as.integer(dStepsPerInterval$interval / 100), formatC(dStepsPerInterval$interval %% 100, width = 2, flag = 0), sep =":")

ggplot(data = dStepsPerInterval, aes(x = interval, y = steps)) + 
    geom_line() + 
    ggtitle("average steps through the day")
```
## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
dStepsPerInterval[which.max(dStepsPerInterval$steps), c("interval2", "steps")]
```

#Q3Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.


##Q3.1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(!complete.cases(dActivity))

sum(is.na(dActivity$date))
sum(is.na(dActivity$interval))
sum(is.na(dActivity$steps))
```
##Q3.2 Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
##Q3.3 Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
dMeanStepsPerInterval   <- aggregate(steps~interval, data = dActivity, FUN = mean)

dActivityNoNA <- dActivity
dActivityNoNA <- merge (dActivityNoNA, dMeanStepsPerInterval, by = "interval")
dActivityNoNA$steps.x <- ifelse (is.na(dActivityNoNA$steps.x), dActivityNoNA$steps.y, dActivityNoNA$steps.x)

dActivityNoNA <- dActivityNoNA[order(dActivityNoNA$date), ]
dActivityNoNA <- subset(dActivityNoNA, select = -c(steps.y))
colnames(dActivityNoNA)[2] <- "steps"
dActivityNoNA<- dActivityNoNA[,c(2,3,1,4)]
```

##Q3.4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
dStepsPerDayNoNA <- aggregate(steps~date, data = dActivityNoNA, FUN = sum, na.rm = TRUE)
barplot(dStepsPerDayNoNA$steps, 
        names.arg = dStepsPerDayNoNA$date, 
        main = "total step per day", 
        ylab = "# steps", 
        xlab = "date")

dMeanStepsPerDayNoNA   <- aggregate(steps~date, data = dActivityNoNA, FUN = mean)
dMedianStepsPerDayNoNA <- aggregate(steps~date, data = dActivityNoNA, FUN = median)
dMeanMedStepsPerDayNoNA <- merge(dMeanStepsPerDayNoNA, dMedianStepsPerDayNoNA, by = "date")
colnames(dMeanMedStepsPerDayNoNA) <- c("date", "mean", "median")
dMeanMedStepsPerDayNoNA
```

#Q4 Are there differences in activity patterns between weekdays and weekends?

##Q4.1 Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
dActivityNoNA$date <- as.POSIXlt(dActivityNoNA$date)
dActivityNoNA$isWeekend <- dActivityNoNA$date$wday >=6
```

##Q4.2 Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
dStepsPerIntervalWeekDayvsWeekens <- aggregate(steps~interval + isWeekend, data = dActivityNoNA, FUN = mean, na.rm = TRUE)
ggplot(data = dStepsPerIntervalWeekDayvsWeekens, aes(x = interval, y = steps, color = isWeekend)) +
    geom_line() +  
    ggtitle("average steps through the day: weekdday vs. weekend")
```